---
title: "facetwarp"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{facetwarp}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

*facetwarp* is an extension of the [*ggplot2*](https://ggplot2.tidyverse.org) specifically aimed at arranging faceted plots.

The main function within the `facetwarp` package is `facet_warp`, which is a close sibling of [ggplot2::facet_wrap](https://ggplot2.tidyverse.org/reference/facet_wrap.html), hence the similar name. 
If you're not already familiar with how to use [ggplot2::facet_wrap](https://ggplot2.tidyverse.org/reference/facet_wrap.html), please start there.

Before you go any further, you should already be familiar with This allows you to 'speak' a graph from composable elements, instead of being limited to a predefined set of charts.


## Part 1: wrap vs warp 🪄

### wrap

First let's recall what `facet_wrap` gives us using the `iris` dataset. 👇
```{r setup}
library(ggplot2)

ggplot(iris) +
  geom_point(aes(x=Petal.Width, y=Petal.Length))+
  facet_wrap(vars(Species),nrow = 2)
```

Note that there are 3 facets, one for each species, and they are arranged in alphabetical order.
Because we've arranged them into 2 rows, and there are only 3 facets, the 4th panel (lower, right) is not occupied.

### warp
Now, we know that we've got other columns in this dataset, specifically `Sepal.Width` and `Sepal.Length`.
Let's explore those axis quickly by summarizing their values by `Species`

```{r}
library(dplyr, warn.conflicts = FALSE)

ggplot(iris %>% 
            group_by(Species) %>% 
            summarize(median_Sepal.Width  = median(Sepal.Width),
                      median_Sepal.Length = median(Sepal.Length)))+
    geom_text(aes(x=median_Sepal.Width, y=median_Sepal.Length, label=Species))
```

In our facetted scatter plot above, instead of arranging the facets alphabetically, 
maybe we want the layout to mimic this `Sepal.Length` and `Sepal.Width` arrangement we see above.

**IT IS TIME TO ✨ WARP THE FACETS  🪄 **


```{r}
library(facetwarp)
ggplot(iris)+
    geom_point(aes(x=Petal.Width, y=Petal.Length))+
    facet_warp(vars(Species), macro_x='Sepal.Width', macro_y='Sepal.Length', nrow = 3, ncol = 3)
```

👆 Notice the layout has changed. `facet_warp` has repositioned the facets!
In fact, they are mimicing the arrangement we saw above:
* `virginica` at the top due to its high `median_Sepal.Length`
* `versicolor` at the left due to its low `median_Sepal.Width`
* `setosa` at the lower-right due to its low `median_Sepal.Length` and high `median_Sepal.Width`

This was accomplished using our **macro** axes.
When we say `macro_x='Sepal.Width'`, we're saying, make no change to the "x axis" of the individual facets, but in order to arrange the facets themselves, treat `Sepal.Width` as the x-dimension.

Since we only need 4 panels total, we can try dropping `nrow` and `ncol` to `2` to condense the arrangement:
```{r}
ggplot(iris)+
    geom_point(aes(x=Petal.Width, y=Petal.Length))+
    facet_warp(vars(Species), macro_x='Sepal.Width', macro_y='Sepal.Length', nrow = 2, ncol = 2)
```


## Part 2: Building on the Warp Idea with Election Data

### The More Facets, The Bigger the Insights

### Leaving Empty Space


## Part 3: Chicago Transity Authority Data
